---
title: "Correlogram RSV CHIPRP-01"
author: "Paloma Garcia"
date: "2026-07-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Correlogram code: install packages


```{r cars}
install.packages("ggcorrplot")
install.packages("ggplot2")
install.packages(c("Hmisc", "corrplot"))
```

## Load packages


```{r pressure, echo=FALSE}
library(readxl)

library(Hmisc)

library(corrplot)

library(dplyr)
```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Read in file 
```{r}
my_data <- read_excel("~/Desktop/Chiu Lab MSc Thesis/Correlogram_PG_older_adults.xlsx")
```

## Check data
```{r}
head(my_data)

str(my_data)
```
## Keep numeric columns only
```{r}
numeric_data <- my_data[sapply(my_data, is.numeric)]
```

## Stop if fewer than 2 numeric columns
```{r}
if (ncol(numeric_data) < 2) {

  stop("Need at least 2 numeric columns for correlation analysis.")

}
```

## Spearman correlation and p-values
```{r}
res2 <- rcorr(as.matrix(numeric_data), type = "spearman")

```

## Optional: quick basic plot
```{r}
corrplot(res2$r)
```

## Adjust p-values using FDR
```{r}
n <- ncol(numeric_data)

Padjust <- matrix(p.adjust(as.vector(res2$P), method = "fdr"),

                  nrow = n, ncol = n)

 

rownames(Padjust) <- rownames(res2$r)

colnames(Padjust) <- colnames(res2$r)
```

## Final correlation plot
```{r}

png(filename = "PG_correlation.png", width = 1200, height = 1200, res = 150)
corrplot(res2$r,
         order = 'hclust',
         addrect = 2,
         col = colorRampPalette(c("#5DA5DA", "white", "#E07A7A"))(200),
         p.mat = Padjust,
         insig = "label_sig",
         sig.level = c(0.001, 0.01, 0.05),
         pch.cex = 0.9,
         pch.col = "black",
         diag = FALSE,
         tl.srt = 90,
         tl.col = "black",
         tl.cex = 0.8,
         title = "Viral Load & Symptoms Correlation with Analytes",
         mar = c(1, 1, 1, 1),
         cl.cex = 0.8
      )
dev.off()

```

